---
title: "ASM_Homerwork_Ex2"
author: "Jamie Arjona Martinez"
date: "September 23, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r imports}
library(tidyverse)
library(pastecs)

```

#ASM

## Exercice 2: Perform the analysis of the CPU time data set that appears at the end of the slides Mirisession2.pdf. Follow the steps of the data analysis stablished as a guidelines.

Seven programs were monitored to observe their resource demands. In particular, the number of disk I/=’s, menory size (in Kilobytes), and CPU time (in milliseconds) were observed. One is interested in modelling CPU time (Y ) as a function of the other two.

```{r cars}
cpu_time <- c(2,5,7,9,10,13,20)
disk <- c(14,16,27,42,39,50,83)
memory_size <- c(70,75,144,190,210,235,400)
data <- data.frame("CPU_TIME" = cpu_time, "DISK" = disk, "MEMORY_SIZE"=memory_size)
```

### Exploratory data analysis
```{r eda}
#Small data set, we can show the whole data
data

#column types
sapply(data, class)

#Descriptive statistics
summary(data)
round(stat.desc(data),2)
```

The size of our data sample is of 7 observations and 3 variables. The dependen variable is CPU_TIME while the independent are DISK i/o and MEMORY_SIZE. It is observed that the units of measure of MEMORY_SIZE is bigger than the other variables and maybe it is worth using normalized data in order to do the modelling, also we observe a discrepance in the range and variance etween the variables.
It is important to note that there are no missing values or NA, so imputations or missing data treatments are not needed.

```{r zscores}
#Normalizing the data
(means <-apply(data, 2, mean))
(sds <- apply(data, 2, sd))
(standardized_data <- as.data.frame(apply(data, 2, function(x) (x-mean(x))/sd(x))))

round(stat.desc(standardized_data),2)
```

With normalized data the units of measure of the variables behave similar and will be easier to detect outliers.

#### Plotting
Boxplots can help to visualize and detect outliers.

```{r plots_no_standardized}
#Boxplot
boxplot(data[, 2:3])
#outlier detected
data[data$MEMORY_SIZE>300,]
```

```{r plots_standardized}
#Boxplot
boxplot(standardized_data[, 2:3])
#outlier detected
standardized_data[standardized_data$MEMORY_SIZE>1.5,]
```

Clearly the program number 7 used more disk and memory that was expected and will have an impact in the modelling. As we don't know the reasons because the program number 7 got this values (computer problems? algorithm behaviour?) we will perform the analysis with and without the program number 7 and see the differences.

```{r remove_7}
data_no_outliers <- data %>% filter(MEMORY_SIZE<300)
standardized_data_no_outliers <- standardized_data %>% filter(MEMORY_SIZE<1.5)
#Boxplot
boxplot(data_no_outliers[, 2:3])
boxplot(standardized_data_no_outliers[, 2:3])
```
Without the program number 7 it seems that we have the data in reasonable ranges.
Next we are going to do the scatterplot of the data.

```{r}
ggplot(data = data) +
  geom_point(mapping = aes(x = DISK, y =cpu_time))

ggplot(data = data) +
  geom_point(mapping = aes(x = memory_size, y =cpu_time))
```
From the scatterplots it is reasonable to expect linear behaviours and so a linear model will fit into the data.

### Hypothesis

### Model

### Testing the hypothesis using the model

### Parameter interpretation

### Answers and Final comments