---
title: "ASM_Homework_Ex3"
output: pdf_document
---

```{r}
#Loading necessary packages
library(car)
library(pastecs)
#library(kableExtra)
#library(tinytex)
library(carData)

```

```{r}
#Loading the data
data <- read.csv("FIFADataset.csv", header = TRUE)

#Filtering the 3 clubs of interest and the necessary variables for the analysis
fc <- data[ data$Club %in% c('Manchester City','Real Madrid CF', 'FC Barcelona'),
            c('Name', 'Agility', 'Acceleration', 'Club', 'Sprint.speed', 'Reactions', 'Balance') ]

#Showing first rows of the data
head(fc)
```
# Description of the Dataset

As it can be seen in the table the data consists of different football players from the FIFA Dataset obtained by Kaggle. This analysis covers only players from the three clubs "Manchester City", "FC Barcelona" and "Real Madrid CF". The following variables are relevant:

* Name: factor, contains the name of the player
* Club: factor, contains name of the club to which a player belongs
* Acceleration: numeric, theoretical range: 1-100
* Agility: numeric, theoretical range: 1-100
* Sprint.speed: numeric, theoretical range: 1-100
* Reactions: numeric, theoretical range: 1-100
* Balance: numeric, theoretical range: 1-100

A more in detail discussion of the variables can be find in the Univariate Descriptive Analysis section.

# Data Cleansing
```{r}
#checking levels
sapply(fc, class)

#looking at the different variables which should be numerical
levels(fc$Acceleration)[fc$Acceleration]
levels(fc$Agility)[fc$Agility]
levels(fc$Sprint.speed)[fc$Sprint.speed]
levels(fc$Reactions)[fc$Reactions]
levels(fc$Balance)[fc$Balance]

#truncate at the first two digits and convert to numerical
fc[, 'Acceleration'] <-as.numeric(substr(levels(fc$Acceleration)[fc$Acceleration], 0, 2))
fc[,'Agility'] <- as.numeric(substr(levels(fc$Agility)[fc$Agility], 0, 2))
fc[, 'Sprint.speed'] <-as.numeric(substr(levels(fc$Sprint.speed)[fc$Sprint.speed], 0, 2))
fc[, 'Reactions'] <-as.numeric(substr(levels(fc$Reactions)[fc$Reactions], 0, 2))
fc[, 'Balance'] <-as.numeric(substr(levels(fc$Balance)[fc$Balance], 0, 2))

#checking levels again
sapply(fc, class)
```

After taking a first look at the data, we see that some values do not consist only out of one number but of some additional numbers which should be truncated, so that all rows of the specific columns can be numeric.

# Univariate Descriptive Analysis and Outlier Detection

```{r}
round(stat.desc(fc[,c('Acceleration', 'Agility', 'Sprint.speed', 'Balance', 'Reactions')], basic = T, desc = T), 2)
```

By looking at the statistics of the numeric variables we see that there are no missing values (neither NA nor NULL) in the data. All variables are more or less in the same range. Only 'Balance' and 'Reactions' have a slightly higher Minimum. Also the means are all around 71 where 'Balance' has the lowest mean (of 69.68) and 'Reactions' the highest mean (of 75.81). The variance values among the variables are similar with the exception of 'Reactions' which has a much smaller variance of 129. 

```{r}
boxplot(fc[, c(2, 3, 5, 6, 7)])
```

It can be seen that there are some outliers in the Acceleration, Sprint.speed and Reactions variables. For futher analysis we will have a closer look on the instances that are smaller than Q1 - 1.5*IQR.

```{r}
fc[fc$Acceleration < 40, ]
fc[fc$Sprint.speed < 40, ]
fc[fc$Reactions < 52, ]
```

The outliers in the Acceleration variable are all goalkeepers. Being a goalkeeper requests completely different abilities than other playing positions as goalkeepers do not have to run through the entire field. As they are a very special case in the Fifa Dataset it is reasonable to exclude all goalkeepers from the analysis. They play under completely different conditions than the other players. However, other players which have very low values e.g. in Sprint.speed or Reactions are not goalkeepers and although they have values which are much lower than the average ones they should remain in the dataset as they play under equal conditions as the other players. 

```{r}
pos <- data[data$Club %in% c('Manchester City','Real Madrid CF', 'FC Barcelona'),
            c('Name', 'Preferred.Positions') ]
#merge the positions to the dataframe so that it is possible to exclude goalkeepers
fc_pos <- merge(pos, fc, by="Name")

#noticed that one player (Danilo from Man City) appeared twice in the original dataset -> remove duplicates
fc_pos <- fc_pos[!duplicated(fc_pos$Name),]

#exclude goalkeepers and remove position variable
fc <- fc_pos[!(fc_pos$Preferred.Positions %in% 'GK '),]
fc <- fc[,c(1, 3, 4, 5, 6, 7, 8)]
```

# Bivariate Descriptive Anaylsis 

```{r}
# Scatter plots
par(mfrow = c(2,2))
plot(fc$Acceleration ~ fc$Agility, ylab = 'Acceleration', xlab='Agility')
plot(fc$Acceleration ~ fc$Sprint.speed, ylab = 'Acceleration', xlab='Sprint.speed')
plot(fc$Acceleration ~ fc$Reactions, ylab = 'Acceleration', xlab='Reactions')
plot(fc$Acceleration ~ fc$Balance, ylab = 'Acceleration', xlab='Balance')

cor(fc[, 'Acceleration'], fc[, 'Agility'])
cor(fc[, 'Acceleration'], fc[, 'Sprint.speed'])
cor(fc[, 'Acceleration'], fc[, 'Reactions'])
cor(fc[, 'Acceleration'], fc[, 'Balance'])
```

In the scatter plots you can see that some variables are more correlated with Acceleration than others. Agility (cor=0.77) and Sprint.speed (cor=0.90) are highly correlated with Acceleration. On the other side the relation in the scatter plots for Reactions (cor=0.36) and Balance (cor=0.57) with Acceleration are not really good to see and the correlation values are low. On a bivariate level therefore Agility and Sprint.Speed should be better predictors for Acceleration than Reactions and Balance. 

# Hypotheses

After the uni and bivariate analysis of the variables the following hypotheses can be stated: 

* The higher the Agility of a player, the higher the Acceleration.
* The higher the Sprint.Speed of a player, the higher the Acceleration.
* The higher the Reactions of a player, the higher the Acceleration.
* The higher the Balance of a player, the higher the Acceleration.
* Agility and Sprint.Speed have a higher influence than Reactions and Balance.

# Linear Model

To check the hypotheses a multivariate linear regression will be performed. 

```{r}
mod <- lm(Acceleration ~ Agility+Sprint.speed+Reactions+Balance, fc)
summary(mod)
```

```{r}
#checking without insignificant variables from first model
mod2 <- lm(Acceleration ~ Agility+Sprint.speed, fc)
summary(mod2)
```







